The sound of teaching music: Expert pianists' performance modulations for novices

### 2022-08-31 paper/manuscript.Rmd
upload analysis for teaching experience
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

# Instructions

@@

# Temporal Variability

@@

# Temporal grouping

@@

# Additional analysis for dynamics difference

@@

>>>>

# 1. Instructions

@@

# 2. Temporal Variability

@@

# 3. Temporal grouping

@@

# 4. Additional analysis for dynamics difference

@@

# 5. Descriptive statistics depending on participants' teaching experience

## Experiment 1
```{r teaching-1, include = FALSE}
source(here("experiment-1/analysis/teaching", "teaching_ioi.R"), chdir = TRUE)
source(here("experiment-1/analysis/teaching", "teaching_art.R"), chdir = TRUE)
source(here("experiment-1/analysis/teaching", "teaching_dyn.R"), chdir = TRUE)
```

```{r ioi-teaching-1, include = FALSE}
# articulation
ioi_art_plot$teachingexp <- factor(ioi_art_plot$teachingexp, levels = c("Yes", "No"))

ioi_art_teaching_1 <- ggpaired(ioi_art_plot, cond1 = "performing", cond2 = "teaching", color = "condition", facet.by = "teachingexp", panel.labs = list(teachingexp = c("Teaching: Yes", "Teaching: No")), palette = "aaas", line.size = 0.3, line.color = "gray", xlab = "Condition", ylab = "IOIs (ms)", title = "IOIs: Articulation") + geom_hline(yintercept = 188, linetype = "dashed") +
  ylim(150, 260) +
  theme_pubr(base_size = 20, legend = "none", base_family = "Helvetica")

# save plots
ggsave(here("paper/image", "ioi_art_teaching_1.png"), plot = ioi_art_teaching_1, dpi = 300, width = 7, height = 5)
img_ioi_art_teaching_1 <- image_read(here("paper/image", "ioi_art_teaching_1.png"), density = 300)

# dynamics
ioi_dyn_plot$teachingexp <- factor(ioi_dyn_plot$teachingexp, levels = c("Yes", "No"))

ioi_dyn_teaching_1 <- ggpaired(ioi_dyn_plot, cond1 = "performing", cond2 = "teaching", color = "condition", facet.by = "teachingexp", panel.labs = list(teachingexp = c("Teaching: Yes", "Teaching: No")), palette = "aaas", line.size = 0.3, line.color = "gray", xlab = "Condition", ylab = "IOIs (ms)", title = "IOIs: Dynamics") + geom_hline(yintercept = 188, linetype = "dashed") +
  ylim(150, 260) +
  theme_pubr(base_size = 20, legend = "none", base_family = "Helvetica")

# save plots
ggsave(here("paper/image", "ioi_dyn_teaching_1.png"), plot = ioi_dyn_teaching_1, dpi = 300, width = 7, height = 5)
img_ioi_dyn_teaching_1 <- image_read(here("paper/image", "ioi_dyn_teaching_1.png"), density = 300)
```

```{r art-teaching-1, include = FALSE}
kot_art_plot$teachingexp <- factor(kot_art_plot$teachingexp, levels = c("Yes", "No"))

art_leg_teaching_1 <- ggpaired(kot_art_plot[subcomponent == "Legato"], cond = "performing", cond2 = "teaching", color = "condition", facet.by = "teachingexp", panel.labs = list(teachingexp = c("Teaching: Yes", "Teaching: No")), palette = "aaas", line.size = 0.3, line.color = "gray", xlab = "Condition", ylab = "KOT (ms)", title = "KOT: Articulation-Legato") + theme_pubr(base_size = 20, legend = "none", base_family = "Helvetica")

# save plots
ggsave(here("paper/image", "art_leg_teaching_1.png"), plot = art_leg_teaching_1, dpi = 300, width = 7, height = 5)
img_art_leg_teaching_1 <- image_read(here("paper/image", "art_leg_teaching_1.png"), density = 300)

art_sta_teaching_1 <- ggpaired(kot_art_plot[subcomponent == "Staccato"], cond = "performing", cond2 = "teaching", color = "condition", facet.by = "teachingexp", panel.labs = list(teachingexp = c("Teaching: Yes", "Teaching: No")), palette = "aaas", line.size = 0.3, line.color = "gray", xlab = "Condition", ylab = "KOT (ms)", title = "KOT: Articulation-Staccato") + theme_pubr(base_size = 20, legend = "none", base_family = "Helvetica")

# save plots
ggsave(here("paper/image", "art_sta_teaching_1.png"), plot = art_sta_teaching_1, dpi = 300, width = 7, height = 5)
img_art_sta_teaching_1 <- image_read(here("paper/image", "art_sta_teaching_1.png"), density = 300)
```

```{r dyn-teaching-1, include = FALSE}
vel_dyn_plot$teachingexp <- factor(vel_dyn_plot$teachingexp, levels = c("Yes", "No"))

vel_for_teaching_1 <- ggpaired(vel_dyn_plot[subcomponent == "Forte"], cond = "performing", cond2 = "teaching", color = "condition", facet.by = "teachingexp", panel.labs = list(teachingexp = c("Teaching: Yes", "Teaching: No")), palette = "aaas", line.size = 0.3, line.color = "gray", xlab = "Condition", ylab = "KV (0-127)", title = "KV: Dynamics-Forte") + theme_pubr(base_size = 20, legend = "none", base_family = "Helvetica")

# save plots
ggsave(here("paper/image", "vel_for_teaching_1.png"), plot = vel_for_teaching_1, dpi = 300, width = 7, height = 5)
img_vel_for_teaching_1 <- image_read(here("paper/image", "vel_for_teaching_1.png"), density = 300)

vel_pia_teaching_1 <- ggpaired(vel_dyn_plot[subcomponent == "Piano"], cond = "performing", cond2 = "teaching", color = "condition", facet.by = "teachingexp", panel.labs = list(teachingexp = c("Teaching: Yes", "Teaching: No")), palette = "aaas", line.size = 0.3, line.color = "gray", xlab = "Condition", ylab = "KV (0-127)", title = "KV: Dynamics-Piano") + theme_pubr(base_size = 20, legend = "none", base_family = "Helvetica")

# save plots
ggsave(here("paper/image", "vel_pia_teaching_1.png"), plot = vel_pia_teaching_1, dpi = 300, width = 7, height = 5)
img_vel_pia_teaching_1 <- image_read(here("paper/image", "vel_pia_teaching_1.png"), density = 300)
```

```{r dyn-diff-teaching-1, include = FALSE}
vel_diff_dyn_plot$teachingexp <- factor(vel_diff_dyn_plot$teachingexp, levels = c("Yes", "No"))

vel_diff_for_teaching_1 <- ggpaired(vel_diff_dyn_plot[subcomponent == "FtoP"], cond = "performing", cond2 = "teaching", color = "condition", facet.by = "teachingexp", panel.labs = list(teachingexp = c("Teaching: Yes", "Teaching: No")), palette = "aaas", line.size = 0.3, line.color = "gray", xlab = "Condition", ylab = "Difference", title = "KV-Diff: Dynamics-FtoP") + theme_pubr(base_size = 20, legend = "none", base_family = "Helvetica")

# save plots
ggsave(here("paper/image", "vel_diff_for_teaching_1.png"), plot = vel_diff_for_teaching_1, dpi = 300, width = 7, height = 5)
img_vel_diff_for_teaching_1 <- image_read(here("paper/image", "vel_diff_for_teaching_1.png"), density = 300)

vel_diff_pia_teaching_1 <- ggpaired(vel_diff_dyn_plot[subcomponent == "PtoF"], cond = "performing", cond2 = "teaching", color = "condition", facet.by = "teachingexp", panel.labs = list(teachingexp = c("Teaching: Yes", "Teaching: No")), palette = "aaas", line.size = 0.3, line.color = "gray", xlab = "Condition", ylab = "Difference", title = "KV-Diff: Dynamics-PtoF") + theme_pubr(base_size = 20, legend = "none", base_family = "Helvetica")

# save plots
ggsave(here("paper/image", "vel_diff_pia_teaching_1.png"), plot = vel_diff_pia_teaching_1, dpi = 300, width = 7, height = 5)
img_vel_diff_pia_teaching_1 <- image_read(here("paper/image", "vel_diff_pia_teaching_1.png"), density = 300)
```

```{r plot-ioi-teaching-1, out.width = "100%", fig.cap = "\\label{fig:ioi-teaching-1}BAA"}
img_ioi_teaching_1 <- image_append(c(img_ioi_art_teaching_1, img_ioi_dyn_teaching_1))
image_write(img_ioi_teaching_1, path = here("paper/figure", "Fig13.tiff"), format = "tiff")
img_ioi_teaching_1
```

```{r plot-art-teaching-1, out.width = "100%", fig.cap = "\\label{fig:art-teaching-1}BAA"}
img_art_teaching_1 <- image_append(c(img_art_leg_teaching_1, img_art_sta_teaching_1))
image_write(img_art_teaching_1, path = here("paper/figure", "Fig14.tiff"), format = "tiff")
img_art_teaching_1
```

```{r plot-dyn-teaching-1, out.width = "100%", fig.cap = "\\label{fig:dyn-teaching-1}BAA"}
img_dyn_teaching_1 <- image_append(c(img_vel_for_teaching_1, img_vel_pia_teaching_1))
image_write(img_dyn_teaching_1, path = here("paper/figure", "Fig15.tiff"), format = "tiff")
img_dyn_teaching_1
```

```{r plot-dyn-diff-teaching-1, out.width = "100%", fig.cap = "\\label{fig:dyn-diff-teaching-1}BAA"}
img_dyn_diff_teaching_1 <- image_append(c(img_vel_diff_for_teaching_1, img_vel_diff_pia_teaching_1))
image_write(img_dyn_diff_teaching_1, path = here("paper/figure", "Fig16.tiff"), format = "tiff")
img_dyn_diff_teaching_1
```

```{r ioi-teaching-desc-1, echo = FALSE}
ioi_all <- rbind(ioi_art_all, ioi_dyn_all)
colnames(ioi_all)[2] <- "Technique"
ioi_all[Condition == "teaching"]$Condition <- "Teaching"
ioi_all[Condition == "performing"]$Condition <- "Performing"
ioi_all[Technique == "articulation"]$Technique <- "Articulation"
ioi_all[Technique == "dynamics"]$Technique <- "Dynamics"

apa_table(
  ioi_all
  , caption = "Caption"
  , note = "Note"
  , escape = TRUE
)
```

```{r art-teaching-desc-1, echo = FALSE}
colnames(kot_art_all)[2] <- "Technique"
kot_art_all[Condition == "teaching"]$Condition <- "Teaching"
kot_art_all[Condition == "performing"]$Condition <- "Performing"
kot_art_all[Technique == "articulation"]$Technique <- "Articulation"

apa_table(
  kot_art_all
  , caption = "Caption"
  , note = "Note"
  , escape = TRUE
)
```

```{r dyn-teaching-desc-1}
colnames(vel_dyn_all)[2] <- "Technique"
vel_dyn_all[Condition == "teaching"]$Condition <- "Teaching"
vel_dyn_all[Condition == "performing"]$Condition <- "Performing"
vel_dyn_all[Technique == "dynamics"]$Technique <- "Dynamics"

apa_table(
  vel_dyn_all
  , caption = "Caption"
  , note = "Note"
  , escape = TRUE
)
```

```{r dyn-diff-teaching-desc-1}
colnames(vel_diff_dyn_all)[2] <- "Technique"
vel_diff_dyn_all[Condition == "teaching"]$Condition <- "Teaching"
vel_diff_dyn_all[Condition == "performing"]$Condition <- "Performing"
vel_diff_dyn_all[Technique == "dynamics"]$Technique <- "Dynamics"

apa_table(
  vel_diff_dyn_all
  , caption = "Caption"
  , note = "Note"
  , escape = TRUE
)
```

## Experiment 2
```{r teaching-2, include = FALSE}
source(here("experiment-2/analysis/teaching", "teaching_ioi.R"), chdir = TRUE)
source(here("experiment-2/analysis/teaching", "teaching_art.R"), chdir = TRUE)
source(here("experiment-2/analysis/teaching", "teaching_dyn.R"), chdir = TRUE)
```

```{r ioi-teaching-2, include = FALSE}
# articulation
ioi_art_plot$teachingexp <- factor(ioi_art_plot$teachingexp, levels = c("Yes", "No"))

ioi_art_teaching_2 <- ggpaired(ioi_art_plot, cond1 = "performing", cond2 = "teaching", color = "condition", facet.by = "teachingexp", panel.labs = list(teachingexp = c("Teaching: Yes", "Teaching: No")), palette = "aaas", line.size = 0.3, line.color = "gray", xlab = "Condition", ylab = "IOIs (ms)", title = "IOIs: Articulation") +
  ylim(0.8, 1.1) +
  geom_hline(yintercept = 1, linetype = "dashed") +
  theme_pubr(base_size = 20, legend = "none", base_family = "Helvetica")

# save plots
ggsave(here("paper/image", "ioi_art_teaching_2.png"), plot = ioi_art_teaching_2, dpi = 300, width = 7, height = 5)
img_ioi_art_teaching_2 <- image_read(here("paper/image", "ioi_art_teaching_2.png"), density = 300)

# dynamics
ioi_dyn_plot$teachingexp <- factor(ioi_dyn_plot$teachingexp, levels = c("Yes", "No"))

ioi_dyn_teaching_2 <- ggpaired(ioi_dyn_plot, cond1 = "performing", cond2 = "teaching", color = "condition", facet.by = "teachingexp", panel.labs = list(teachingexp = c("Teaching: Yes", "Teaching: No")), palette = "aaas", line.size = 0.3, line.color = "gray", xlab = "Condition", ylab = "IOIs (ms)", title = "IOIs: Dynamics") +
  ylim(0.8, 1.1) +
  geom_hline(yintercept = 1, linetype = "dashed") +
  theme_pubr(base_size = 20, legend = "none", base_family = "Helvetica")

# save plots
ggsave(here("paper/image", "ioi_dyn_teaching_2.png"), plot = ioi_dyn_teaching_2, dpi = 300, width = 7, height = 5)
img_ioi_dyn_teaching_2 <- image_read(here("paper/image", "ioi_dyn_teaching_2.png"), density = 300)
```

```{r art-teaching-2, include = FALSE}
kor_art_plot$teachingexp <- factor(kor_art_plot$teachingexp, levels = c("Yes", "No"))

art_leg_teaching_2 <- ggpaired(kor_art_plot[subcomponent == "Legato"], cond = "performing", cond2 = "teaching", color = "condition", facet.by = "teachingexp", panel.labs = list(teachingexp = c("Teaching: Yes", "Teaching: No")), palette = "aaas", line.size = 0.3, line.color = "gray", xlab = "Condition", ylab = "KOR", title = "KOR: Articulation-Legato") + theme_pubr(base_size = 20, legend = "none", base_family = "Helvetica")

# save plots
ggsave(here("paper/image", "art_leg_teaching_2.png"), plot = art_leg_teaching_2, dpi = 300, width = 7, height = 5)
img_art_leg_teaching_2 <- image_read(here("paper/image", "art_leg_teaching_2.png"), density = 300)

art_sta_teaching_2 <- ggpaired(kor_art_plot[subcomponent == "Staccato"], cond = "performing", cond2 = "teaching", color = "condition", facet.by = "teachingexp", panel.labs = list(teachingexp = c("Teaching: Yes", "Teaching: No")), palette = "aaas", line.size = 0.3, line.color = "gray", xlab = "Condition", ylab = "KOR", title = "KOR: Articulation-Staccato") + theme_pubr(base_size = 20, legend = "none", base_family = "Helvetica")

# save plots
ggsave(here("paper/image", "art_sta_teaching_2.png"), plot = art_sta_teaching_2, dpi = 300, width = 7, height = 5)
img_art_sta_teaching_2 <- image_read(here("paper/image", "art_sta_teaching_2.png"), density = 300)
```

```{r dyn-teaching-2, include = FALSE}
vel_dyn_plot$teachingexp <- factor(vel_dyn_plot$teachingexp, levels = c("Yes", "No"))

vel_for_teaching_2 <- ggpaired(vel_dyn_plot[subcomponent == "Forte"], cond = "performing", cond2 = "teaching", color = "condition", facet.by = "teachingexp", panel.labs = list(teachingexp = c("Teaching: Yes", "Teaching: No")), palette = "aaas", line.size = 0.3, line.color = "gray", xlab = "Condition", ylab = "KV (0-127)", title = "KV: Dynamics-Forte") + theme_pubr(base_size = 20, legend = "none", base_family = "Helvetica")

# save plots
ggsave(here("paper/image", "vel_for_teaching_2.png"), plot = vel_for_teaching_2, dpi = 300, width = 7, height = 5)
img_vel_for_teaching_2 <- image_read(here("paper/image", "vel_for_teaching_2.png"), density = 300)

vel_pia_teaching_2 <- ggpaired(vel_dyn_plot[subcomponent == "Piano"], cond = "performing", cond2 = "teaching", color = "condition", facet.by = "teachingexp", panel.labs = list(teachingexp = c("Teaching: Yes", "Teaching: No")), palette = "aaas", line.size = 0.3, line.color = "gray", xlab = "Condition", ylab = "KV (0-127)", title = "KV: Dynamics-Piano") + theme_pubr(base_size = 20, legend = "none", base_family = "Helvetica")

# save plots
ggsave(here("paper/image", "vel_pia_teaching_2.png"), plot = vel_pia_teaching_2, dpi = 300, width = 7, height = 5)
img_vel_pia_teaching_2 <- image_read(here("paper/image", "vel_pia_teaching_2.png"), density = 300)
```

```{r dyn-diff-teaching-2, include = FALSE}
vel_diff_dyn_plot$teachingexp <- factor(vel_diff_dyn_plot$teachingexp, levels = c("Yes", "No"))

vel_diff_for_teaching_2 <- ggpaired(vel_diff_dyn_plot[subcomponent == "FtoP"], cond = "performing", cond2 = "teaching", color = "condition", facet.by = "teachingexp", panel.labs = list(teachingexp = c("Teaching: Yes", "Teaching: No")), palette = "aaas", line.size = 0.3, line.color = "gray", xlab = "Condition", ylab = "Difference", title = "KV-Diff: Dynamics-FtoP") + theme_pubr(base_size = 20, legend = "none", base_family = "Helvetica")

# save plots
ggsave(here("paper/image", "vel_diff_for_teaching_2.png"), plot = vel_diff_for_teaching_2, dpi = 300, width = 7, height = 5)
img_vel_diff_for_teaching_2 <- image_read(here("paper/image", "vel_diff_for_teaching_2.png"), density = 300)

vel_diff_pia_teaching_2 <- ggpaired(vel_diff_dyn_plot[subcomponent == "PtoF"], cond = "performing", cond2 = "teaching", color = "condition", facet.by = "teachingexp", panel.labs = list(teachingexp = c("Teaching: Yes", "Teaching: No")), palette = "aaas", line.size = 0.3, line.color = "gray", xlab = "Condition", ylab = "Difference", title = "KV-Diff: Dynamics-PtoF") + theme_pubr(base_size = 20, legend = "none", base_family = "Helvetica")

# save plots
ggsave(here("paper/image", "vel_diff_pia_teaching_2.png"), plot = vel_diff_pia_teaching_2, dpi = 300, width = 7, height = 5)
img_vel_diff_pia_teaching_2 <- image_read(here("paper/image", "vel_diff_pia_teaching_2.png"), density = 300)
```

```{r plot-ioi-teaching-2, out.width = "100%", fig.cap = "\\label{fig:ioi-teaching-2}BAA"}
img_ioi_teaching_2 <- image_append(c(img_ioi_art_teaching_2, img_ioi_dyn_teaching_2))
image_write(img_ioi_teaching_2, path = here("paper/figure", "Fig17.tiff"), format = "tiff")
img_ioi_teaching_2
```

```{r plot-art-teaching-2, out.width = "100%", fig.cap = "\\label{fig:art-teaching-2}BAA"}
img_art_teaching_2 <- image_append(c(img_art_leg_teaching_2, img_art_sta_teaching_2))
image_write(img_art_teaching_2, path = here("paper/figure", "Fig18.tiff"), format = "tiff")
img_art_teaching_2
```

```{r plot-dyn-teaching-2, out.width = "100%", fig.cap = "\\label{fig:dyn-teaching-2}BAA"}
img_dyn_teaching_2 <- image_append(c(img_vel_for_teaching_2, img_vel_pia_teaching_2))
image_write(img_dyn_teaching_2, path = here("paper/figure", "Fig19.tiff"), format = "tiff")
img_dyn_teaching_2
```

```{r plot-dyn-diff-teaching-2, out.width = "100%", fig.cap = "\\label{fig:dyn-diff-teaching-2}BAA"}
img_dyn_diff_teaching_2 <- image_append(c(img_vel_diff_for_teaching_2, img_vel_diff_pia_teaching_2))
image_write(img_dyn_diff_teaching_2, path = here("paper/figure", "Fig20.tiff"), format = "tiff")
img_dyn_diff_teaching_2
```

```{r ioi-teaching-desc-2, echo = FALSE}
ioi_all <- rbind(ioi_art_all, ioi_dyn_all)
colnames(ioi_all)[2] <- "Technique"
ioi_all[Condition == "teaching"]$Condition <- "Teaching"
ioi_all[Condition == "performing"]$Condition <- "Performing"
ioi_all[Technique == "articulation"]$Technique <- "Articulation"
ioi_all[Technique == "dynamics"]$Technique <- "Dynamics"
ioi_all$Teaching <- factor(ioi_all$Teaching, levels = c("Yes", "No"))
setorder(ioi_all, "Technique", "Teaching")

apa_table(
  ioi_all
  , caption = "Caption"
  , note = "Note"
  , escape = TRUE
  , digits = 3
)
```

```{r art-teaching-desc-2, echo = FALSE}
colnames(kor_art_all)[2] <- "Technique"
kor_art_all[Condition == "teaching"]$Condition <- "Teaching"
kor_art_all[Condition == "performing"]$Condition <- "Performing"
kor_art_all[Technique == "articulation"]$Technique <- "Articulation"
kor_art_all$Teaching <- factor(kor_art_all$Teaching, levels = c("Yes", "No"))
setorder(kor_art_all, "Technique", "Teaching")

apa_table(
  kor_art_all
  , caption = "Caption"
  , note = "Note"
  , escape = TRUE
  , digits = 3
)
```

```{r dyn-teaching-desc-2}
colnames(vel_dyn_all)[2] <- "Technique"
vel_dyn_all[Condition == "teaching"]$Condition <- "Teaching"
vel_dyn_all[Condition == "performing"]$Condition <- "Performing"
vel_dyn_all[Technique == "dynamics"]$Technique <- "Dynamics"
vel_dyn_all$Teaching <- factor(vel_dyn_all$Teaching, levels = c("Yes", "No"))
setorder(vel_dyn_all, "Technique", "Teaching")

apa_table(
  vel_dyn_all
  , caption = "Caption"
  , note = "Note"
  , escape = TRUE
)
```

```{r dyn-diff-teaching-desc-2}
colnames(vel_diff_dyn_all)[2] <- "Technique"
vel_diff_dyn_all[Condition == "teaching"]$Condition <- "Teaching"
vel_diff_dyn_all[Condition == "performing"]$Condition <- "Performing"
vel_diff_dyn_all[Technique == "dynamics"]$Technique <- "Dynamics"
vel_diff_dyn_all$Teaching <- factor(vel_diff_dyn_all$Teaching, levels = c("Yes", "No"))
setorder(vel_diff_dyn_all, "Technique", "Teaching")

apa_table(
  vel_diff_dyn_all
  , caption = "Caption"
  , note = "Note"
  , escape = TRUE
)
```
